Procedural generation of levels for the angry birds videogame using evolutionary computation\\

### 2019-05-14 IEEE_ConferenceOnGames/article_english.tex
Introduction edited

\section{Introduction}
Procedural content generation is an interesting subject on the entertainment
industry, primarily in the video game development because of the easiness and
quickness provided to generate content for different video game projects [ref] .
In this paper an evolutionary content generation system is proposed to generate
levels for the popular Angry Birds video game
\cite{RovioEntertainmentCorporation2009}, taking inspiration in concepts of
open-ended evolution.

For this project a procedural generation content system is proposed for the game
angry birds, angry birds is a game developed by Rovio Entertainment in which the
objective is to destroy structures placed on the map as well as eliminating
green pig-like entities in order to obtain the highest score possible, the game
uses a gravity and collition system that allows the structures to fall and be
destroyed by means of damage as well as the use of three different construction
materials \textit{ice}, \textit{wood} and \textit{stone} each one with different
resistances ranging from fragile, normal and sturdy in that order, the game
mechanic is to use a slingshot to shoot a certain number of birds with different
abilities to destroy the structures and the objectives, it is possible to obtain
a high score by doing the most destruction as possible with the least birds as
you can. \cite{RovioEntertainmentCorporation2009}

While it is possible to create a evolutionary agent to play the levels and
obtain the best score the main objective of this project is to generate the
levels that can be played by naive agents or physical players, the levels
themselves need to be complex in the sense that interesting structures can be
created but it has to be possible to complete in some way
\cite{Stephenson,Stephenson2018}.

% What is the problem?
Procedural content generation provides a peculir case of study because of the way
content has to be made for a each particular type of game, in this particular case 
the restrictions that the game and the competition track have provide an interesting 
challenge for a evolutionary algorithm because we have to tanke into account the 
different pieces that can be used, their interactions with one another because of 
gravity and the possible effects the trajectory of a shot from a player can affect 
all of this.

% How we solve it? (summary)   
In order to tackle this case of study we proposed to use a genetic algorithm capable 
of creating content, more specifically levels for the angry birds video game in which
the generated levels satisfy the objectives of being \textit{interesting} and 
\textit{playable}.

@@

As mentioned above in this paper we use procedural content generation (PCG) 
which denotes the way to automatically generate
content by algorithms instead of generating the same content by manual means
such as with different personnel.

In the videogame industry many developers use a PCG style software that produces
raw images and designs used by other personnel to improve it and use it in order
to reduce the development cycles, and the productions cost. 
PCG as various focus areas that resolve around a video game, however in this paper 
the main focus is the \textit{Level design} area which as the objective to design 
the different areas of a game in which a player can interact with, since in the 
angry birds video game a level is represented by a scene then the system will 
have to create the entities (birds and pigs) and objects (pieces) in it in order 
to create a playable level.

And finally we use our own interpretation of an Open-ended evolution (OEE) 
algorithm which is a variant of genetic evolution algorithm where
the objective differs from the regular genetic algorithm one, in this case the
final objective is not to reach a final state of the evolution but to continue
evolving and creating new types of entities more complex on each new generation,
these same entities have the possibility to evolve not only inside the frame of
their type but create entirely new groups. \cite{Standish2003}

Another explanation provided by T. Taylor et al. \cite{Taylor2016,Taylor}
defines OEE has a system capable of creating novelty beyond a point where
nothing changes on the current group of interest rather than just converging to
a stable or quasi-stable state.

For this project OEE is defined simply has a process of genetic evolution that
encapsulates a group of similar processes and that is able to increase complexity
 of the original entities.

@@

Section 2, contains a quick overview of some aspects used by different authors
in order to tackle this problem, their design and basic functionality, in
Section 3 the description of the current state of the project is explained, from
the basis of the development, modification to the original proposed method as 
well as different approaches that were taken to solve the problem.
% to the different methods that have been used to
%progress on the different sections of the project, 
% Normaly there is no "Future Work" section is put together with Conclusions - MG
finally in Section 4 we provide our conclusions about the development of this
paper as well as improvements that can help to refine the proposed work.

>>>>

\section{Introduction}
Procedural content generation is an interesting subject on the entertainment
industry, primarily in video game development because of the savings in 
time and human resources when generating content for different video game projects.
In this paper an evolutionary content generation system is proposed to generate
levels for the popular Angry Birds video game \cite{RovioEntertainmentCorporation2009},
taking inspiration in concepts of open-ended evolution. Angry birds is a game 
developed by Rovio Entertainment in which the objective is to destroy structures
placed on the map as well as eliminating
green pig-like entities in order to obtain the highest score possible, the game
uses a gravity and collition system that allows the structures to fall and be
destroyed by means of damage as well as the use of three different construction
materials \textit{ice}, \textit{wood} and \textit{stone} each one with a different
resistance, ranging from fragile, normal and sturdy in that order. The game
mechanic is to use a slingshot to shoot a certain number of birds with different
abilities, to destroy both the structures and the objectives. A high score 
is achived by doing the most destruction with the least birds possible 
\cite{RovioEntertainmentCorporation2009}.

%While it is possible to create an evolutionary agent to play the levels and
%obtain the best score 
The main objective of this project is to generate levels that can be played
by naïve or inteligent agents as well as physical players. The levels
themselves need to be complex in the sense that interesting structures can be
created, but at the same time, it is possible to complete in some way
\cite{Stephenson,Stephenson2018}.

% What is the problem?
Procedural content generation provides a peculiar case of study because of the way
% Is there a reference? - MG
content has to be made for a each particular type of game. In this use case 
we have those restrictions dictated by the game, and also 
those imposed by a competition track \cite{Renz}.% Reference the paper of the competition  have provide an interesting 
The challenge for the participating algorithms is to be flexible on the  
pieces that can be used and the runing time of the algorithm, as they are determined at runtime. 
%their interactions with one another because of 
%gravity and the possible effects the trajectory of a shot from a player can affect 
%all of this.
% How we solve it? (summary)   
In order to tackle this case study, we proposed the use a genetic algorithm capable 
of creating game levels having the objective of being both \textit{interesting} and 
\textit{playable}.

@@

% Not important - MG
%As mentioned above in this paper we use procedural content generation (PCG) 
%which denotes the way to automatically generate
%content by algorithms instead of generating the same content by manual means
%such as with different personnel.

%In the videogame industry many developers use a PCG style software that produces
%raw images and designs used by other personnel to improve it and use it in order
%to reduce the development cycles, and the productions cost. 
%PCG as various focus areas that resolve around a video game, however in this paper 
%the main focus is the \textit{Level design} area which as the objective to design 
%the different areas of a game in which a player can interact with, since in the 
%angry birds video game a level is represented by a scene then the system will 
%have to create the entities (birds and pigs) and objects (pieces) in it in order 
%to create a playable level.

We propose a basic Open-Ended Evolution (OEE) algorithm, a variant of genetic evolution algorithm where
the objective differs from that of a regular genetic algorithm,  in this case the
final objective is not to reach a final state of the evolution (an optimal solution) but to continue
evolving and creating new types of entities, perhaps more complex on each new generation.
Solutuion entities have the possibility to evolve not only inside the frame of
their type, but  they can create entirely new groups \cite{Standish2003}. T. Taylor et al. \cite{Taylor2016,Taylor}
defines OEE as a system capable of creating novelty beyond a point where
nothing changes on the current group of interest rather than just converging to
a stable or quasi-stable state.

For this project, OEE is defined simply as a genetic evolution process with
different clasess of objects, with compositional and taxonomic relationships between them.
Objects of each class can evolve in parallel or as in this case, independenly of each other.   
The idea is based on the object oriented model, in which complex behavior can be model 
using these principles.

@@

Section 2, contains a quick overview of some aspects used by different authors
in order to tackle this problem, their design and basic functionality, in
Section 3 the description of the current state of the project is explained, from
%% The following sentence is not clear.
the basis of the development, modification to the original proposed method as 
well as different approaches that were taken to solve the problem.
% to the different methods that have been used to
%progress on the different sections of the project, 
% Normaly there is no "Future Work" section is put together with Conclusions - MG
finally in Section 4 we provide our conclusions about the development of this
paper as well as improvements that can help to refine the proposed work.
